---
title: "Assessment_1"
author: "Bailey Dalton"
date: "2024-11-11"
output: html_document
---

#### Section 1: Loading Libraries

```{r Chunk 1, message = FALSE, warning = FALSE, eval = TRUE, echo = FALSE}
library(tidyverse)
library(knitr)
library(dplyr)
```

```{r LoadDataset, eval = TRUE, echo = TRUE}

# Read in dataset 
LifeExp_data <- read.csv("data/life-expectancy-un-vs-gdp-per-capita-wb.csv")

# Data cleanup and wrangling to ensure data fits correct specifications for assessment 1
LifeExp_data <- LifeExp_data %>%
  # rename columns to simplify
  rename(
    Country = Entity,
    Life_expectancy = Life.expectancy...Sex..all...Age..0...Variant..estimates,
    GDP_per_capita = GDP.per.capita..PPP..constant.2017.international...,
    Population = Population..historical.
  ) %>% 
  # Select only relevant variables 
  select(Country, Year, Life_expectancy, GDP_per_capita, Population) %>% 
  # Remove missing values 
  filter(!is.na(GDP_per_capita) & !is.na(Life_expectancy) & !is.na(Population)) %>% 
  # Filter to keep four years (in 5-year intervals), reducing rows to < 1000
  filter(Year %in% c(2005, 2010, 2015, 2020))
  

glimpse(LifeExp_data)
```


#### Section 2: Research Question 

This report will explore the relationship between GDP per capita, a key indicator of a country's wealth, and its correlation with life expectancy and population across multiple countries from 2005 to 2020. The analysis aims to determine whether GDP directly influences both life expectancy and population size during this period.


#### Section 3: Data Set Introduction





#### Section 4: Data Set Description 
